---
title: "Plots and Code"
output: 
  html_document:
    toc: true
    toc_float: true
    smooth_scroll: true
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```



### Time Intervals {.tabset}

Patterns that occur over a 24-hour time frame can be tricky to communicate effectively, especially when major transitions in those trends happen at night. There are several different ways to represent time in R; here's just a couple of potential solutions.

#### Circular Plots

```{r}
# simulate data

# activity for 6 species, 24 hours
data = data.frame(hour = rep(seq(1,24),6),
                  species = rep(c("Cow","Elephant","Zebra","Giraffe","Hyena","Lion"), each=24))

# activity a function of time
data$activity = c(dnorm(data$hour[which(data$species=="Cow")], mean=12, sd=2), 
                  dnorm(data$hour[which(data$species=="Elephant")], mean=14, sd=5),
                  dnorm(data$hour[which(data$species=="Zebra")], mean=10, sd=3),
                  dnorm(data$hour[which(data$species=="Giraffe")], mean=15, sd=1),
                  (dnorm(data$hour[which(data$species=="Hyena")], mean=4, sd=2)+
                    dnorm(data$hour[which(data$species=="Hyena")], mean=20, sd=2))/2,
                  (dnorm(data$hour[which(data$species=="Lion")], mean=2, sd=2)+
                    dnorm(data$hour[which(data$species=="Lion")], mean=24, sd=2))/2)


# activity is a probability in this case, but it could also be a count etc.
head(data)

```


Now let's plot these data using ggplot2:

```{r}

ggplot(data, aes(x=hour, y=activity))+
  geom_bar(aes(fill=species, color=species), stat="identity", alpha=0.5) +
  theme_bw(base_size=20)+
  coord_polar()


```

Interesting! So here, we can see some variation in animal activity. However, it's a little challenging to compare, say carnivores and herbivores. Let's play with some of the aesthetics:

```{r}

# I usually define colors that I want using the color picker option:
levels(as.factor(data$species))
# go to Addins > Colour Picker and choose based on the levels. I like to return color names

animalcols = c("darkseagreen", "aquamarine4", "darkolivegreen3", "tomato2", "firebrick", "lightgoldenrod3")

circ_plot = ggplot(data, aes(x=hour, y=activity))+
  geom_rect(aes(xmin=0,xmax=6,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_rect(aes(xmin=18,xmax=24,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_col(aes(fill=species, color=species),position=position_nudge(-0.5),alpha=0.3,width=1) +
  theme_bw(base_size=20)+
  scale_fill_manual(values=animalcols)+
  scale_color_manual(values=animalcols)+
  scale_x_continuous(breaks=c(0,6,12,18))+
  theme(axis.text.x=element_text(vjust=1))+
  annotate(geom="text", x=c(0,12), y=c(0.3,0.3), label=c("Night","Day"), size=5)+
  coord_polar(theta="x", clip="off")

circ_plot
```


If desired, you can also explore faceting by species:

```{r}
ggplot(data, aes(x=hour, y=activity))+
  geom_rect(aes(xmin=0,xmax=6,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_rect(aes(xmin=18,xmax=24,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_col(aes(fill=species, color=species),position=position_nudge(-0.5),alpha=0.3,width=1) +
  theme_bw(base_size=20)+
  scale_fill_manual(values=animalcols)+
  scale_color_manual(values=animalcols)+
  scale_x_continuous(breaks=c(0,6,12,18))+
  theme(axis.text.x=element_text(vjust=1))+
    coord_polar(theta="x", clip="off")+
  facet_wrap(~species) 
```

Unfortunately, free x and y scales are not supported for circular plots, so data transformation might be necessary to see these effectively.


#### Non-circular Options

Another option is to stack the data:

```{r}
ggplot(data, aes(x=hour, y=activity))+
  geom_rect(aes(xmin=0,xmax=6,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_rect(aes(xmin=18,xmax=24,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_col(aes(fill=species, color=species), alpha=0.5) +
  theme_bw(base_size=20)+
  scale_fill_manual(values=animalcols)+
  scale_color_manual(values=animalcols)+
  coord_cartesian(xlim=c(0,24))+
  scale_x_continuous(breaks=c(0,6,12,18,24))+
  annotate(geom="text", x=c(2,12), y=c(0.6,0.6), label=c("Night","Day"), size=5)

# overlapping
ggplot(data, aes(x=hour, y=activity))+
  geom_rect(aes(xmin=0,xmax=6,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_rect(aes(xmin=18,xmax=24,ymin=-Inf,ymax=Inf),fill="gray70",alpha=0.1)+
  geom_col(aes(fill=species, color=species), position=position_nudge(-0.5),alpha=0.3) +
  theme_bw(base_size=20)+
  scale_fill_manual(values=animalcols)+
  scale_color_manual(values=animalcols)+
  coord_cartesian(xlim=c(0,24))+
  scale_x_continuous(breaks=c(0,6,12,18,24))+
  annotate(geom="text", x=c(2,12), y=c(0.6,0.6), label=c("Night","Day"), size=5)

```


